---
title: "Data Preparation - HDB"
author: "Daniel Ang"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading packages
```{r message=FALSE, warning=FALSE}
packages = c('sf', 'dplyr', 'tidyverse', 'readr', 'httr', 'kableExtra', 'tmap', 'rgeos', 'rgdal', 'plotKML')
for(p in packages){
  if(!require(p, character.only =T)){
    install.packages(p)
  }
  library(p,character.only =T)
}
memory.size(max = FALSE)
memory.limit(size = NA)

print_table = function(x){
  n = nrow(x)
  if (n > 6){n = 6}
  x[1:n,] %>%
    kable(format = "html", escape = F) %>%
    kable_styling(bootstrap_options = c("striped", "bordered", "responsive"))
}
```

# Importing data

Importing resale HDB pricing
```{r}
resalePrices = read_csv(file="Raw/HDB resale/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv",
                        col_types = cols(
                          month = col_date(format = "%Y-%m"),
                          town = col_factor(),
                          flat_type = col_factor(),
                          block = col_character(),
                          street_name = col_character(),
                          storey_range = col_factor(),
                          floor_area_sqm = col_double(),
                          flat_model = col_factor(),
                          lease_commence_date = col_date(format = "%Y"),
                          remaining_lease = col_character(),
                          resale_price = col_double()
                        ))

```

Feature creation to get `price_per_sqm`, `age_month`, `remaining_lease_month`

```{r}
# Transforming "x year x month" in lease_commence_date into just "x month"
getMonths = function(remaining_lease){
  year_month = regmatches(remaining_lease, gregexpr('\\(?[0-9,.]+', remaining_lease))
  year = as.numeric(year_month[[1]][1])
  month = as.numeric(year_month[[1]][2])
  if(is.na(year)){year = 0}
  if(is.na(month)){month = 0}
  return(year*12 + month)
}

# Get months difference between 2 different date value
elapsed_months = function(end_date, start_date) {
  ed <- as.POSIXlt(end_date)
  sd <- as.POSIXlt(start_date)
  12 * (ed$year - sd$year) + (ed$mon - sd$mon)
}


resalePrices = resalePrices %>% 
  rowwise() %>% 
  mutate("transaction_year" = format(`month`,"%Y")) %>%
  mutate("price_per_sqm" = round(resale_price/floor_area_sqm, digits=2)) %>%
  mutate("age_month" = elapsed_months(`month`, `lease_commence_date`)) %>%
  mutate("remaining_lease_month" = getMonths(`remaining_lease`)) %>%
  select(-c("lease_commence_date", "remaining_lease", "month"))

print_table(resalePrices)
```

# Geolocation search
## Creating list of unique address to perform geolocation search

Creating a list of unique search address for geo-location search
```{r}
uniqueAddress = select(resalePrices, `street_name`, `block`) %>%
  unique()
```

Searching unique address up on oneMap API
```{r}
# HTTP get query to OneMap API 
search_address_info = function(street_name, block) {
  search_query = paste(`street_name`, `block`, sep=" ")
  search_query = gsub(" ", "%20", search_query)
  
  r = RETRY("GET", paste("https://developers.onemap.sg/commonapi/search?searchVal=",search_query,"&returnGeom=Y&getAddrDetails=Y", sep = ""))
  response_text = content(r, "parsed")
  if(response_text$found != 0){
    return(response_text$results[1])
  }else{
    return(NA)
  }
}

uniqueAddress = uniqueAddress %>% 
  rowwise() %>% 
  mutate("result" = search_address_info(`street_name`, `block`))
```


Unloading results from OneMap API
```{r}
uniqueAddress = uniqueAddress %>% 
  mutate("postal" = result[6][[1]]) %>% 
  mutate("address" = result[5][[1]]) %>% 
  mutate("latitude" = result[9][[1]]) %>% 
  mutate("longtitude" = result[10][[1]]) %>%
  select(-"result")
```

At this point, we now have postal code, lat, long for each of our HDB address
```{r}
print_table(uniqueAddress)
head(uniqueAddress)
```

# Count of facilities for each postal code

## Importing masterplan subzone data

```{R}
mpsz = st_read(dsn = "raw/MP14_SUBZONE", layer = "MP14_SUBZONE_WEB_PL")
mpsz_svy21 = st_transform(mpsz, 3414)
```

## Creating a sf object from the aspatial data frame
```{r}
# getting rid of any failed variables
uniqueAddress = uniqueAddress %>% 
  filter(!is.na(latitude))

# Creating a sf object from the aspatial data frame
uniqueAddress.sf = sf::st_as_sf(uniqueAddress, coords = c("longtitude", "latitude"), crs=4326) %>% 
  st_transform(crs=3414)
```

Setting distance buffer
```{r}
# distance in meters
buffersize = 1500
```

## Shopping mall
```{R} 
countWithinBuffer.shp = function(geometry, locationList.shp, buffersize){
  pointsBuffer = gBuffer(as(geometry, "Spatial"), width = buffersize)
  proj4string(locationList.shp) = CRS("+init=epsg:3414")
  proj4string(pointsBuffer) = CRS("+init=epsg:3414")
  
  count = sp::over(locationList.shp, pointsBuffer)
  c = as.data.frame(count) %>% filter(!is.na(count))
  return(nrow(c))
}

# loading shopping mall data
list_of_shopping_malls = read.csv(file="Processed/Shopping malls/list_of_shopping_malls.csv", header=TRUE, sep=",")
list_of_shopping_malls = list_of_shopping_malls %>% 
  filter(!is.na(latitude))

list_of_shopping_malls.sf = st_as_sf(list_of_shopping_malls, coords = c("longtitude", "latitude"), crs=4326) %>% 
  st_transform(crs=3414)

# Visualizing bus interchange locations
plot(st_geometry(mpsz), main="Shopping mall")
plot(st_geometry(list_of_shopping_malls.sf), col = "red", lwd = 3, add = TRUE)

# Counting qty of malls within buffer
list_of_shopping_malls.shp = as(list_of_shopping_malls.sf, 'Spatial')
uniqueAddress.sf = uniqueAddress.sf %>%
  rowwise() %>%
  mutate("mallCount" = countWithinBuffer.shp(`geometry`, list_of_shopping_malls.shp, buffersize))

head(uniqueAddress.sf)
```
## Bus Interchange
```{r}
# loading bus interchange data
list_of_bus_interchange = read.csv(file="Processed/Bus interchange/list_of_bus_interchange.csv", header=TRUE, sep=",")
list_of_bus_interchange = list_of_bus_interchange %>% 
  filter(!is.na(latitude))

list_of_bus_interchange.sf = st_as_sf(list_of_bus_interchange, coords = c("longtitude", "latitude"), crs=4326) %>% 
  st_transform(crs=3414)

# Visualizing bus interchange locations
plot(st_geometry(mpsz), main="Bus Interchange")
plot(st_geometry(list_of_bus_interchange.sf), col = "red", lwd = 3, add = TRUE)


# Counting qty of malls within buffer
list_of_bus_interchange.shp = as(list_of_bus_interchange.sf, 'Spatial')
uniqueAddress.sf = uniqueAddress.sf %>%   
  rowwise() %>%
  mutate("busInterchangeCount" = countWithinBuffer.shp(`geometry`, list_of_bus_interchange.shp, buffersize))

head(uniqueAddress.sf)
```


## SportSG Sport Facilities
```{R}
countWithinBuffer.kml = function(geometry, locationList.kml, buffersize){
  pointsBuffer = gBuffer(as(geometry, "Spatial"), width=buffersize)
  st_crs(locationList.kml) = st_crs(pointsBuffer)
  
  count = sp::over(as(locationList.kml, "Spatial"), pointsBuffer)
  c = as.data.frame(count) %>% filter(!is.na(count))
  return(nrow(c))
}


sportsg_sport_facilities = st_read("raw/SportSG Sport Facilities/sportsg-sport-facilities-kml.kml") %>% st_transform(crs=3414)
sportsg_sport_facilities = st_zm(sportsg_sport_facilities, drop = TRUE, what = "ZM")

# Visualizing SportSG Sport Facilities locations
plot(st_geometry(mpsz), main="SportSG Sport Facilities")
plot(st_geometry(sportsg_sport_facilities), col = "red", lwd = 3, add = TRUE)

uniqueAddress.sf = uniqueAddress.sf %>%
  rowwise() %>%
  mutate("sportsCount" = countWithinBuffer.kml(`geometry`, sportsg_sport_facilities, buffersize))

head(uniqueAddress.sf)
```


## Hawker
```{R}
hawker_facilities = st_read("raw/Hawker centres/hawker-centres-kml.kml") %>% st_transform(crs=3414)
hawker_facilities = st_zm(sportsg_sport_facilities, drop = TRUE, what = "ZM")

# Visualizing Hawker Facilities locations
plot(st_geometry(mpsz), main="Hawker Facilities")
plot(st_geometry(hawker_facilities), col = "red", lwd = 3, add = TRUE)

uniqueAddress.sf = uniqueAddress.sf %>%
  rowwise() %>%
  mutate("hawkerCount" = countWithinBuffer.kml(`geometry`, hawker_facilities, buffersize))

head(uniqueAddress.sf)
```

## NParks Parks
```{R}
nparklocation = st_read("raw/NParks Parks/nparks-parks-kml.kml") %>% st_transform(crs=3414)
nparklocation_pt = st_centroid(nparklocation)$geometry

# Visualizing Hawker Facilities locations
plot(st_geometry(mpsz), main="NParks Parks")
plot(st_geometry(nparklocation_pt), col = "red", lwd = 3, add = TRUE)

uniqueAddress.sf = uniqueAddress.sf %>%
  rowwise() %>%
  mutate("npParkCount" = countWithinBuffer.kml(`geometry`, nparklocation_pt, 500))

head(uniqueAddress.sf)
```

## Pre-school locations
```{r}
preschoolslocation = st_read("raw/List of Preschools/pre-schools-location-kml.kml") %>% st_transform(crs=3414)

# Visualizing Pre-school locations
plot(st_geometry(mpsz), main="Pre-school locations")
plot(st_geometry(preschoolslocation), col = "red", lwd = 3, add = TRUE)

uniqueAddress.sf = uniqueAddress.sf %>%
  rowwise() %>%
  mutate("preschoolCount" = countWithinBuffer.kml(`geometry`, preschoolslocation, buffersize))

head(uniqueAddress.sf)
```

## School location
```{r}
# loading school location data
list_of_school = read.csv(file="Processed/School Directory and Information/general-information-of-schools.csv", header=TRUE, sep=",")
list_of_school = list_of_school %>% 
  filter(!is.na(latitude))

list_of_school.sf = st_as_sf(list_of_school, coords = c("longtitude", "latitude"), crs=4326) %>% 
  st_transform(crs=3414)

# Visualizing bus interchange locations
plot(st_geometry(mpsz), main="School location")
plot(st_geometry(list_of_school.sf), col = "red", lwd = 3, add = TRUE)

list_of_school.sf.primary = list_of_school.sf %>% filter(mainlevel_code == "PRIMARY")
list_of_school.sf.secondary = list_of_school.sf %>% filter(mainlevel_code == "SECONDARY")
list_of_school.sf.tertiary = list_of_school.sf %>% filter(!mainlevel_code %in% c('PRIMARY','SECONDARY'))

# Counting qty of malls within buffer
list_of_bus_interchange.shp.primary = as(list_of_school.sf.primary, 'Spatial')
list_of_bus_interchange.shp.secondary = as(list_of_school.sf.secondary, 'Spatial')
list_of_bus_interchange.shp.tertiary = as(list_of_school.sf.tertiary, 'Spatial')


uniqueAddress.sf = uniqueAddress.sf %>%   
  rowwise() %>%
  mutate("primarySchCount" = countWithinBuffer.shp(`geometry`, list_of_bus_interchange.shp.primary, buffersize)) %>%
  mutate("secondarySchCount" = countWithinBuffer.shp(`geometry`, list_of_bus_interchange.shp.secondary, buffersize)) %>%
  mutate("tertiarySchCount" = countWithinBuffer.shp(`geometry`, list_of_bus_interchange.shp.tertiary, buffersize))

head(uniqueAddress.sf)
```

# Joining data
```{r}
street_to_MRT = read_csv(file="Raw/Distance to MRT/street_to_MRT.csv")
street_to_MRT = street_to_MRT %>% 
  select(-"X1") %>%
  filter(`street_name` %in% uniqueAddress.sf$street_name) %>%
  distinct(`street_name`, .keep_all = TRUE)

# Merging resalePrices & additional postal code info
resalePrices_merged = merge(resalePrices, uniqueAddress.sf, by=c("block","street_name"))

# Merging resalePrices & distance to mrt
resalePrices_merged = merge(resalePrices_merged, street_to_MRT, by=c("street_name"))



# Merging resalePrices & median rental by type
medianRental = read_csv(file="Processed/Median Rent by Town and Flat Type/median-rent-by-town-and-flat-type.csv")
medianRental = medianRental %>% rename("transaction_year" = "year")
resalePrices_merged = left_join(resalePrices_merged, medianRental, by=c("transaction_year", "town", "flat_type"))

# Export the final file
write_csv(x = resalePrices_merged, path = "Processed/resalePrices_merged.csv")
```